\LARGE \bf Human-Robot Interaction: Tackling the AI Challenges

### 2014-05-06 paper.tex
US->GB, typos, capitalization

Human-robot interaction is a challenge for artificial intelligence. This field
lays at the crossroad of several domains of AI and requires to tackle them in a
holistic manner: Modeling humans and human cognition; acquiring, representing,
manipulating in a tractable way abstract knowledge at the human level; reasoning
on this knowledge to make decisions; and eventually instantiating those
decisions into physical actions both legible to and in coordination with humans.
Many AI techniques are invited, from visual processing to symbolic reasoning,
from task planning to \emph{theory of mind} building, from reactive control to
action recognition and learning.

We do not claim to address here the whole issue. This article attempts however to organize
it into a coherent challenge for artificial intelligence, and to explain and
illustrate some of the paths that we have investigated on our robots and that
result in a deliberative, knowledge-oriented, architecture for human-robot
interaction.

We focus on a specific class of interactions: collaborative task achievement
supported by multi-modal and situated communication. Figure~\ref{fig:hri-dec}
illustrates this context: the two agents share a common space and exchange
information through multiple modalities, and the robot is expected to achieve
interactive object manipulation, fetch and carry tasks and other similar
domestic tasks by taking into account, at every stages, the intentions, beliefs,
perspectives, skills of the human partner.  Namely, the robot must be able to
recognize, understand and participate to communication situations, both explicit
(\eg the human addresses verbally the robot) and implicit (\eg the human points
to an object); the robot must be able to take part to joint actions, both
proactively (by planning and proposing resulting plans to the human) and
reactively; the robot must be able to move and act in a safe, efficient and
legible way, taking into account social rules like proxemics.

@@

The remaining of the article discusses the robotic architecture we have built to
tackle the AI challenges of human-robot interaction, and how its relates to
other approaches. We propose to organize this discussion in four sections.

The next section presents the overall architecture and the knowledge model we
have developed for our robots. Section~\ref{sec:impl} gives some insights on
each of the main components of the architecture, and try to highlight their
significance for artificial intelligence. Section~\ref{sec:expe} presents two
experiments that illustrate in a concrete way what can be currently achieved
with our robots.  Section~\ref{sect|conclusion} finally summarizes our main
contributions and relate them to other recent researches in the field.

@@

  %%% SHARY
  \node at (4,4)[skill, fill=RoyalBlue!50] (shary) {%
      \begin{tikzpicture}
        \node at (0,0) (exec) {Execution Controller};
        \node [subpart, below=0.2 of exec.south west, anchor=north west] (plans) {Goal \& Plans \\ management};
        \node [subpart, right=0.2 of plans] (sit-asses) {Situation assessment \\ and context management};
        \node [subpart, right=0.2 of sit-asses] {Action instanciation, \\ execution and monitoring};
      \end{tikzpicture}
    };

@@

For instance, when processing a sentence like ``give me another book'', the {\sc
Dialogs} module sends at some point a request to the knowledge base that looks
like {\tt \footnotesize find(?book type Book, ?book differentFrom
BOOK1)}\footnote{We present a similar but complete example at
section~\ref{moving-london}.} and may write back assertions like \stmt{HUMAN
desires GIVE\_ACTION45, GIVE\_ACTION45 actsOn BOOK2} to {\sc Oro}. The HATP
planner then uses the knowledge base to initialize the planning domains with
similar requests ({\tt \footnotesize find(BOOK2 isAt ?location)}, etc.). The
execution controller typically monitor conditions (by subscribing to events
like: {\tt \footnotesize onNewMatch(HUMAN desires ?goal)}) and stores what the
robot is currently doing (\stmt{myself currentlyPerforms GIVE\_ACTION45}).

@@

Our architecture however relates to \emph{Beliefs, Desires, Intentions} (BDI)
architectures. As put by Woolridge~\cite{Woolridge1999}, BDI architectures are
primarily focused on \emph{practical reasoning}, \ie the process of deciding,
step by step, which action to perform to reach a goal. The management of the
interaction between knowledge (the beliefs) and task and plan representation and
execution (the desires and the intentions) is central, and aims at selecting at
each step the best subgoal. It becomes then an intention that the robot commits
to.

@@

This architecture design (a central knowledge base that essentially appears as a
passive component to the rest of the system -- even though it actually actively
processes the knowledge pool in the background, for instance with the reasoner)
departs from other approaches like the CAST model~\cite{Hawes2007} where
knowledge is represented as a diffuse resource, or the CRAM/KnowRob
architecture~\cite{Beetz2010} where the knowledge base is an active hub that
pro-actively queries perceptual components to acquire knowledge. We believe that
our design leads to a good \emph{observability} (knowledge flows are explicit
and easy to capture since they are centralized) as well as high modularity
(modules communicate through a defined and unified API).

At run-time, the knowledge available to the robot at a given time comes from
three sources. {\it A priori} knowledge is stored in an ontology (the {\sc
OpenRobots} ontology, discussed below) and loaded at startup. This static source covers the
\emph{common-sense} knowledge of the robot, and optionally some
scenario-specific knowledge (for instance, about objects that are to be
manipulated). The second part of the knowledge is acquired at run-time from
perception, interaction and planning. The next sections detail these processes.
Lastly, the third source of symbolic statements is the reasoner itself.

Contrary to other projects like KnowRob~\cite{Tenorth2009a} that relies on
the concept of \emph{computables} to lazily evaluate/acquire symbolic facts when
needed, we have an \emph{explicit} approach where we greedly compute and assert
symbolic statements (like spatial relations between objects, see
Section~\ref{sect|sit-ass}). This design choice trades scalability for explicit
reasoning: at any time, the full belief state is explicitely stated, and
provides the reasoner with the largest possible inference domain. This is of
special importance for an \emph{event-based} architecture like ours (see
Section~\ref{sect|ctrl}), where, by definition, we do not know beforehand what
we are ``looking for'', and we need instead greedly infer as much as possible
to build a belief state as comprehensive as possible.

@@

The tree in Figure~\ref{fig|upper_tbox}\footnote{This subset of the ontology is
indeed a tree. This has however not to be the case in general, and, as a matter
of fact, the TBox of the whole {\sc Oro} common-sense ontology does not form a tree}
is not equally developed in each directions at lower levels. The descendants of
\concept{PartiallyTangibleThing} (\ie what we commonly call \emph{objects}), for
instance, are pictured in Figure~\ref{fig|tangible_things_tbox}. This excerpt
from the ontology makes the bottom-up design process clear: only few types of
\emph{partially tangible} things appear, and those are only subclasses relevant
to the context of service robotics in an human-like environment. For performance
reasons as well as clarity, we do not attempt to proactively extend the
conceptual coverage of the field. We instead opportunistically extend the
common-sense knowledge when required by experiments.

@@

We call \emph{internal} those cognitive capabilities that are tightly bound to
the knowledge model, and hence implemented directly within the {\sc Oro} server.
We present here three of them: \emph{reasoning}, \emph{theory of mind} modeling
and our (naive) approach to \emph{memory management}.

@@

From a robotics point of view, it supposes the ability to build, store and
retrieve separate models of the beliefs of the agents the robot interacts with.
Our knowledge base implements such a mechanism: when the robots recognises that
a new agent has been created in the knowledge base, it initializes a new,
independent, ontology for this agent. All the ontologies that are created share
the same common-sense knowledge, but rely on each agent's perspective for the
actual instantiation: the robot (geometrically) computes that the book is in its
own field of view, but not in the human one. The robot knowledge contains the
fact \stmt{book isVisible true} while the human model contains \stmt{book
isVisible false} (this computation, called \emph{perspective taking}, is
discussed in the next section).

@@

\begin{itemize}
	\item \concept{isOn}: computes if an object $O_1$ is on another object $O_2$ by
	evaluating the center of mass of $O_1$ according to the bounding box of $O_2$.

@@

\item \emph{Reachable}: it allows the robot to estimate the agent's capability
to reach an object, which is fundamental for task planning. For example, if the
user asks the robot to give him/her an object, the robot must compute a transfer
point where the user is able to get the object afterwards. 
Figure~\ref{fig::reach-ex} explains how addressee-dependent reachability is
computed, based on Generalized Inverse Kinematics and collision detection.

@@

Monitoring human activity is crucial for maintaining a coherent state of the
world. Full human action and activity monitoring is a task that requires
knowledge and reasoning both on high level facts like goals, intentions and
plans, as well as bottom-up data from agent and object motions. Simple temporal
and geometric reasoning on human hand trajectories and potential objects
placements can provide some useful clues for high level human monitoring
processes. We call this temporal and geometric reasoning \emph{primitive action
recognition}. Those primitive actions are assessed through monitoring situations
like ``an empty hand is close to an object on a table'' (precursor for a
\emph{pick}), or ``a hand holding an object is over a container'' (precursor for
a \emph{throw}). \emph{Spark} recognizes a handful \fxfatal{which ones?} of these
primitives, that are relied on by the robot to track the state transitions in
the human plans (as explained below, at section~\ref{hatp}).

@@

\subsection{Human-aware task planning}
\label{hatp}

@@

\paragraph{Agents and action streams}

@@

We also use shared plans to verbalize the sequence of actions and to explain
the human partner how a task may be shared~\cite{warnier2012when}.

@@

\paragraph{Action costs and social rules}

@@

\caption{An alternative plan where the \emph{wasted time} social rule is used to optimize the
total duration of the task. }

@@

\paragraph{Several levels of cooperation} 
By tuning its costs
and adapting its social rules, HATP can be used to compute various
alternative plans. These plans can be categorized into several levels
of cooperation

@@

\subsubsection{Action execution and monitoring}\label{sec:action}

@@

\paragraph{Human aware motion, placement and manipulation planning}
Manipulation and navigation actions listed in
table~\ref{table|pyrobots_actions} rely on a collection of dedicated 3D motion
planner to generate movements that are aware of surrounding humans. Robot
placement and end-effectors trajectories are computed on-line to allow object
manipulation that take into account task specific constraints and human
postures, abilities and preferences. \cite{Sisbot2008, Mainprice2011,
Pandey2011} detail these techniques.

@@

\section{Experimental outcomes}
\label{sec:expe}

Our architecture has been deployed and tested in a number of experiments on
several robotic platforms. Table~\ref{table|experiences} lists the most
significant ones, with their main focuses and reference publications for the
interested reader.

@@

\begin{tabular}{llll}
 \bf{Experiment} & Focus & \emph{Notes} & Reference \\
\hline
{\it Point \& Learn} (2010) & Interactive knowledge acquisition & & \cite{Lemaignan2010} \\
{\it Spy Game} (2010) & Interactive object discrimination & & \cite{Ros2010b} \\
{\it Interactive Grounding I} (2011) & Multi-modal interaction, perspective taking & \emph{immobile robot} & \cite{lemaignan2011what} \\
{\it Roboscopie} (2011) & Theater, reflection on the future of HRI & \emph{semi-autonomous} & \cite{lemaignan2012roboscopie} \\
{\it Cleaning the table} (2011) & Whole stack integration & & \cite{Alami2011} \\
{\it I'm in your shoes} (2012) & False beliefs & & \cite{warnier2012when} \\
{\it Give me this} (2012) & Natural joint object manipulation & & \cite{gharbi2013natural} \\
{\it Interactive Grounding II} (2012) & Multi-modal interaction, perspective taking & & \cite{lemaignan2013talking} \\
\hline

@@

Because only objects marked with tags are currently recognized (in our typical
lab experiments, it may be a dozen of them), the robot grounding mechanisms work
in a small-sized closed world. This assumption obviously greatly facilitates
the task. It remains to be seen if systematic marking of manipulated objects
(for instance with RFID tags) is a realistic path. It may be the case if the
robot is expected to be used in a well-defined, slowly changing environment.
Real-time recognition of large libraries of objects is another path which may
be viable soon~\cite{Dean2013Fast}.

@@

The cognitive model that the robot builds for the agent it interacts with is
still simple and mostly focused on geometric features (who sees what? What are
our relative positions? etc.). Extending this knowledge with more subtle
perceptions (emotional state for instance) remains to be done.

@@

Indeed, we posit that human-robot collaborative task achievement calls for more
subtlety for a robot than autonomous independent task achievement. HATP, by
co-planning for the robot and the human, and modeling social rules as planning
parameters, is a first step in this direction.

\paragraph{Joint actions with humans}\label{sec:soa}
The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasize that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realize it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they
all have shown benefits of this approach. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way. Finally, today only few
systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
account at all levels.

@@

One prototypical example of semantic disambiguation has been given in
\cite{Ros2010b} with the child game \emph{spygame}: two players are facing
each other with a set of random objects in-between, one player mentally choose
one object, and the other player has to guess the object by asking closed
questions like \emph{Is your object small or large?} Based on the knowledge it
has acquired, the robot is able to minimize the number of questions required to
find the object.

@@

\section*{Acknowledgments}

>>>>

Human-robot interaction is a challenge for artificial intelligence. This field
lays at the crossroad of several domains of AI and requires to tackle them in a
holistic manner: Modelling humans and human cognition; acquiring, representing,
manipulating in a tractable way abstract knowledge at the human level; reasoning
on this knowledge to make decisions; and eventually instantiating those
decisions into physical actions both legible to and in coordination with humans.
Many AI techniques are invited, from visual processing to symbolic reasoning,
from task planning to \emph{theory of mind} building, from reactive control to
action recognition and learning.

We do not claim to address here the whole issue. This article attempts however to organise
it into a coherent challenge for artificial intelligence, and to explain and
illustrate some of the paths that we have investigated on our robots and that
result in a deliberative, knowledge-oriented, architecture for human-robot
interaction.

We focus on a specific class of interactions: collaborative task achievement
supported by multi-modal and situated communication. Figure~\ref{fig:hri-dec}
illustrates this context: the two agents share a common space and exchange
information through multiple modalities, and the robot is expected to achieve
interactive object manipulation, fetch and carry tasks and other similar
domestic tasks by taking into account, at every stages, the intentions, beliefs,
perspectives, skills of the human partner.  Namely, the robot must be able to
recognise, understand and participate to communication situations, both explicit
(\eg the human addresses verbally the robot) and implicit (\eg the human points
to an object); the robot must be able to take part to joint actions, both
proactively (by planning and proposing resulting plans to the human) and
reactively; the robot must be able to move and act in a safe, efficient and
legible way, taking into account social rules like proxemics.

@@

The remaining of the article discusses the robotic architecture we have built to
tackle the AI challenges of human-robot interaction, and how its relates to
other approaches. We propose to organise this discussion in four sections.

The next section presents the overall architecture and the knowledge model we
have developed for our robots. Section~\ref{sec:impl} gives some insights on
each of the main components of the architecture, and try to highlight their
significance for artificial intelligence. Section~\ref{sec:expe} presents two
experiments that illustrate in a concrete way what can be currently achieved
with our robots.  Section~\ref{sect|conclusion} finally summarises our main
contributions and relate them to other recent researches in the field.

@@

  %%% SHARY
  \node at (4,4)[skill, fill=RoyalBlue!50] (shary) {%
      \begin{tikzpicture}
        \node at (0,0) (exec) {Execution Controller};
        \node [subpart, below=0.2 of exec.south west, anchor=north west] (plans) {Goal \& Plans \\ management};
        \node [subpart, right=0.2 of plans] (sit-asses) {Situation assessment \\ and context management};
        \node [subpart, right=0.2 of sit-asses] {Action instantiation, \\ execution and monitoring};
      \end{tikzpicture}
    };

@@

For instance, when processing a sentence like ``give me another book'', the {\sc
Dialogs} module sends at some point a request to the knowledge base that looks
like {\tt \footnotesize find(?book type Book, ?book differentFrom
BOOK1)}\footnote{We present a similar but complete example at
section~\ref{moving-london}.} and may write back assertions like \stmt{HUMAN
desires GIVE\_ACTION45, GIVE\_ACTION45 actsOn BOOK2} to {\sc Oro}. The HATP
planner then uses the knowledge base to initialise the planning domains with
similar requests ({\tt \footnotesize find(BOOK2 isAt ?location)}, etc.). The
execution controller typically monitor conditions (by subscribing to events
like: {\tt \footnotesize onNewMatch(HUMAN desires ?goal)}) and stores what the
robot is currently doing (\stmt{myself currentlyPerforms GIVE\_ACTION45}).

@@

Our architecture however relates to \emph{Beliefs, Desires, Intentions} (BDI)
architectures. As put by Woolridge~\cite{Woolridge1999}, BDI architectures are
primarily focused on \emph{practical reasoning}, \ie the process of deciding,
step by step, which action to perform to reach a goal. The management of the
interaction between knowledge (the beliefs) and task and plan representation and
execution (the desires and the intentions) is central, and aims at selecting at
each step the best sub-goal. It becomes then an intention that the robot commits
to.

@@

This architecture design (a central knowledge base that essentially appears as a
passive component to the rest of the system -- even though it actually actively
processes the knowledge pool in the background, for instance with the reasoner)
departs from other approaches like the CAST model~\cite{Hawes2007} where
knowledge is represented as a diffuse resource, or the CRAM/KnowRob
architecture~\cite{Beetz2010} where the knowledge base is an active hub that
pro-actively queries perceptual components to acquire knowledge. We believe that
our design leads to a good \emph{observability} (knowledge flows are explicit
and easy to capture since they are centralised) as well as high modularity
(modules communicate through a defined and unified API).

At run-time, the knowledge available to the robot at a given time comes from
three sources. {\it A priori} knowledge is stored in an ontology (the {\sc
OpenRobots} ontology, discussed below) and loaded at start-up. This static source covers the
\emph{common-sense} knowledge of the robot, and optionally some
scenario-specific knowledge (for instance, about objects that are to be
manipulated). The second part of the knowledge is acquired at run-time from
perception, interaction and planning. The next sections detail these processes.
Lastly, the third source of symbolic statements is the reasoner itself.

Contrary to other projects like KnowRob~\cite{Tenorth2009a} that relies on
the concept of \emph{computables} to lazily evaluate/acquire symbolic facts when
needed, we have an \emph{explicit} approach where we greedily compute and assert
symbolic statements (like spatial relations between objects, see
Section~\ref{sect|sit-ass}). This design choice trades scalability for explicit
reasoning: at any time, the full belief state is explicitly stated, and
provides the reasoner with the largest possible inference domain. This is of
special importance for an \emph{event-based} architecture like ours (see
Section~\ref{sect|ctrl}), where, by definition, we do not know beforehand what
we are ``looking for'', and we need instead greedily infer as much as possible
to build a belief state as comprehensive as possible.

@@

The tree in Figure~\ref{fig|upper_tbox}\footnote{This subset of the ontology is
indeed a tree. This has however not to be the case in general, and, as a matter
of fact, the TBox of the whole {\sc Oro} common-sense ontology does not form a tree}
is not equally developed in each directions at lower levels. The descendants of
\concept{PartiallyTangibleThing} (\ie what we commonly call \emph{objects}), for
instance, are pictured in Figure~\ref{fig|tangible_things_tbox}. This excerpt
from the ontology makes the bottom-up design process clear: only few types of
\emph{partially tangible} things appear, and those are only subclasses relevant
to the context of service robotics in an human-like environment. For performance
reasons as well as clarity, we do not attempt to pro-actively extend the
conceptual coverage of the field. We instead opportunistically extend the
common-sense knowledge when required by experiments.

@@

We call \emph{internal} those cognitive capabilities that are tightly bound to
the knowledge model, and hence implemented directly within the {\sc Oro} server.
We present here three of them: \emph{reasoning}, \emph{theory of mind} modelling
and our (naive) approach to \emph{memory management}.

@@

From a robotics point of view, it supposes the ability to build, store and
retrieve separate models of the beliefs of the agents the robot interacts with.
Our knowledge base implements such a mechanism: when the robots recognises that
a new agent has been created in the knowledge base, it initialises a new,
independent, ontology for this agent. All the ontologies that are created share
the same common-sense knowledge, but rely on each agent's perspective for the
actual instantiation: the robot (geometrically) computes that the book is in its
own field of view, but not in the human one. The robot knowledge contains the
fact \stmt{book isVisible true} while the human model contains \stmt{book
isVisible false} (this computation, called \emph{perspective taking}, is
discussed in the next section).

@@

\begin{itemize}
	\item \concept{isOn}: computes if an object $O_1$ is on another object $O_2$ by
	evaluating the centre of mass of $O_1$ according to the bounding box of $O_2$.

@@

\item \emph{Reachable}: it allows the robot to estimate the agent's capability
to reach an object, which is fundamental for task planning. For example, if the
user asks the robot to give him/her an object, the robot must compute a transfer
point where the user is able to get the object afterwards. 
Figure~\ref{fig::reach-ex} explains how addressee-dependent reachability is
computed, based on Generalised Inverse Kinematics and collision detection.

@@

Monitoring human activity is crucial for maintaining a coherent state of the
world. Full human action and activity monitoring is a task that requires
knowledge and reasoning both on high level facts like goals, intentions and
plans, as well as bottom-up data from agent and object motions. Simple temporal
and geometric reasoning on human hand trajectories and potential objects
placements can provide some useful clues for high level human monitoring
processes. We call this temporal and geometric reasoning \emph{primitive action
recognition}. Those primitive actions are assessed through monitoring situations
like ``an empty hand is close to an object on a table'' (precursor for a
\emph{pick}), or ``a hand holding an object is over a container'' (precursor for
a \emph{throw}). \emph{Spark} recognises a handful \fxfatal{which ones?} of these
primitives, that are relied on by the robot to track the state transitions in
the human plans (as explained below, at section~\ref{hatp}).

@@

\subsection{Human-Aware Task Planning}
\label{hatp}

@@

\subsubsection{Agents and Action Streams}

@@

We also use shared plans to verbalise the sequence of actions and to explain
the human partner how a task may be shared~\cite{warnier2012when}.

@@

\subsubsection{Action Costs and Social Rules}

@@

\caption{An alternative plan where the \emph{wasted time} social rule is used to optimise the
total duration of the task. }

@@

\subsubsection{Commitment Levels}

By tuning its costs and adapting its social rules, HATP can be used to compute
various alternative plans. These plans can be categorised into several levels of
cooperation

@@

\subsubsection{Action Execution and Monitoring}\label{sec:action}

@@

\paragraph{Human-Aware Motion, Placement and Manipulation Planning}
Manipulation and navigation actions listed in
table~\ref{table|pyrobots_actions} rely on a collection of dedicated 3D motion
planner to generate movements that are aware of surrounding humans. Robot
placement and end-effectors trajectories are computed on-line to allow object
manipulation that take into account task specific constraints and human
postures, abilities and preferences. \cite{Sisbot2008, Mainprice2011,
Pandey2011} detail these techniques.

@@

\section{Experimental Support}
\label{sec:expe}

Our architecture has been deployed and tested in a number of studies on
several robotic platforms. Table~\ref{table|experiences} lists the most
significant ones, with their main focuses and reference publications for the
interested reader.

@@

\begin{tabular}{llll}
 \bf{Study} & Focus & \emph{Notes} & Reference \\
\hline
{\it Point \& Learn} (2010) & Interactive knowledge acquisition & & \cite{Lemaignan2010} \\
{\it Spy Game} (2010) & Interactive object discrimination & & \cite{Ros2010b} \\
{\it Interactive Grounding I} (2011) & Multi-modal interaction, perspective taking & \emph{immobile robot} & \cite{lemaignan2011what} \\
{\it Roboscopie} (2011) & Theater, reflection on the future of HRI & \emph{semi-autonomous} & \cite{lemaignan2012roboscopie} \\
{\it Cleaning the table} (2011) & Whole stack integration & & \cite{Alami2011} \\
{\it I'm in your shoes} (2012) & False beliefs & & \cite{warnier2012when} \\
{\it Give me this} (2012) & Natural joint object manipulation & & \cite{gharbi2013natural} \\
{\it Interactive Grounding II} (2012) & Multi-modal interaction, perspective taking & & \cite{lemaignan2013talking} \\
\hline

@@

Because only objects marked with tags are currently recognised (in our typical
lab experiments, it may be a dozen of them), the robot grounding mechanisms work
in a small-sized closed world. This assumption obviously greatly facilitates
the task. It remains to be seen if systematic marking of manipulated objects
(for instance with RFID tags) is a realistic path. It may be the case if the
robot is expected to be used in a well-defined, slowly changing environment.
Real-time recognition of large libraries of objects is another path which may
be viable soon~\cite{Dean2013Fast}.

@@

The cognitive model that the robot builds for the agent it interacts with is
still simple and mostly focused on geometric features (\emph{Who sees what? What are
our relative positions?} etc.). Extending this knowledge with more subtle
perceptions (emotional state for instance) remains to be done.

@@

Indeed, we posit that human-robot collaborative task achievement calls for more
subtlety for a robot than autonomous independent task achievement. HATP, by
co-planning for the robot and the human, and modelling social rules as planning
parameters, is a first step in this direction.

\paragraph{Joint actions with humans}\label{sec:soa}
The human presence brings new requirements for robot's abilities both
at the functional and at the deliberative levels~\cite{Klein2004}. The
topics involve motion~\cite{Kulic2007,Berg2004,Madhava2006},
navigation~\cite{Althaus2004,Sisbot2007}, manipulation~\cite{Kemp2007}
in presence of humans as well as perception of human
activities~\cite{Breazeal2001,Burger2008}. Also, when
interacting with humans, robots need to incorporate communication and
collaboration abilities. Several theories dealing with
collaboration~\cite{Cohen1991,Grosz1996,Clark1996} emphasise that
collaborative tasks have specific requirements compared to individual
ones, \eg, since the robot and the person share a common goal, they
have to agree on the manner to realise it, they must show their
commitment to the goal during execution, etc. Several robotic systems
have already been built based on these
theories~\cite{Rich1997,Sidner2005,Breazeal2003} and they
all have shown benefits of this approach. They have also shown how
difficult it is to manage turn-taking between communication partners
and to interleave task realization and communication in a generic
way. Finally, today only few
systems~\cite{Fong2006,Breazeal2003,Sisbot2008} take humans into
account at all levels.

@@

One prototypical example of semantic disambiguation has been given in
\cite{Ros2010b} with the child game \emph{spygame}: two players are facing
each other with a set of random objects in-between, one player mentally choose
one object, and the other player has to guess the object by asking closed
questions like \emph{Is your object small or large?} Based on the knowledge it
has acquired, the robot is able to minimise the number of questions required to
find the object.

@@

\section*{Acknowledgements}
